---
title: "Untitled"
author: "Oisín McNulty"
date: "20/04/2021"
output: html_document
---



```{r setup, include=FALSE}
## THIS WILL CREATE A VERY LARGE OBJECT (4mb) WHICH WILL FREEZE SMALL MULTICORE COMPUTERS.  To replicate reduce t and m to 10's
t <- 500
m <- 100
h <- array(rnorm(m*t),c(m,t)) # time series in rows
e <- h %*% t(h)/t # form the correlation matrix
lambda_e <- eigen(e,symmetric = T, only.values = T)
ee <- lambda_e$values # Eigenvalyes if the correlation matrix

load("C:/Users/Oisin/Dropbox/My PC (LAPTOP-FFR71B0C)/Downloads/daily_factors.rda")
load("C:/Users/Oisin/Dropbox/My PC (LAPTOP-FFR71B0C)/Downloads/ftse350.rda")
load("C:/Users/Oisin/Dropbox/My PC (LAPTOP-FFR71B0C)/Downloads/ftse25_rtns_mthly.rda")

library(magrittr)

library(fpc)
library(factoextra)
library(fungible)
library(koRpus)
library(tools)

library(entropy)
library(tidyverse)
library(ati)
library(ggdendro)

```



# Marcenko-Pastur density
```{r}
#Function to compute Marcenko-Pastur density in
library(matlab)
mp_pdf<-function(var,t,m,pts) {
  q=t/m
  eMin<-var*(1-(1./q)^.5)^2 
  eMax<-var*(1+(1./q)^.5)^2 
  eVal<-linspace(eMin,eMax,pts)
  pdf<-q/(2*pi*var*eVal)*((eMax-eVal)*(eVal-eMin))^.5
  pdf<-array(pdf) 
  names(pdf)<-eVal # creates a named array
  return(pdf)
}

```


# Random Matrix Technique 
```{r}
custom.portfolio.moments <- function(R, portfolio) {
  momentargs<-list()
  momentargs$mu<-matrix(as.vector(apply(R,2, "mean")), ncol = 1)
  momentargs$sigma<-estRMT(R, parallel=FALSE)$cov
   momentargs$m3 <- matrix(0, nrow=ncol(R), ncol=ncol(R)^2)
  momentargs$m4 <- matrix(0, nrow=ncol(R), ncol=ncol(R)^3)
  return(momentargs)
}

```

```{r}
setwd("C:/Users/Oisin/OneDrive - Queen's University Belfast/Quant - Msc 2020/Semester 2/FIN7030 - Algo trading and investing")
dow30data <- read.csv("Algorithmic Trading and Investment.csv")
```


```{r}
library(PortfolioAnalytics)
pspec.lo <- portfolio.spec(assets = colnames(dow30data))
# Specification 1 and 2
pspec.lo <- add.constraint(pspec.lo, type="full_investment")
pspec.lo <- add.constraint(pspec.lo, type="long_only")
# Specification 3
pspec.lo <- add.objective(portfolio=pspec.lo, type="return", name="mean")
pspec.lo <- add.objective(portfolio=pspec.lo, type="risk", name="var")


```
```{r}

```


# Distance Metrics


```{r}
#Information Entropy - given probabilities
p <- c(0.3,0.7)
-sum(p*log(p))

#Information Entropy - given probabilities
p <- c(0.01,0.99)
-sum(p*log(p))

#Information Entropy - given probabilities
p<-c(0.7,0.15,0.15)
-sum(p*log(p))


```

# NMI Statistics
```{R}
numBins<-function(nObs,corr=NULL){
  #Optimal
  if (is.null(corr)) {
    z=(8+324*nObs+12*(36*nObs+729*nObs^2)^0.5)^(1/3)
    b=round(z/6+2/(3*z)+1/3)
  }
  else {
    #bivariate case
    b=round(2^(-0.5)*(1+(1+24*nObs/(1-corr^2))^0.5)^0.5)
  }
  return(b)
}
```

```{r, echo=TRUE, include=FALSE}
#install.packages("RTransferEntropy")
#install.packages("future")
library(RTransferEntropy)
library(future)
# enable parallel processing
plan(multisession) # initialised a multicore enviroment
data("stocks") # loads data

TE<-stocks %>%
  group_by(ticker) %>%
  group_split(.keep = TRUE) %>%
  map(~transfer_entropy(x = .x$ret,y=.x$sp500,shuffles = 500,type = "bins",bins=12))
names(TE)<-unique(stocks$ticker)
```

```{r, fig.show = "hold", out.width="50%"}
ent_table <- TE$coef
for (i in 1:length(TE)) {
    ent_table$names[i] <- names(TE)[i]
    ent_table$te[i] <- TE[[i]][["coef"]][1,1]
    ent_table$ete[i] <- TE[[i]][["coef"]][1,2]
    ent_table$se[i] <- TE[[i]][["coef"]][1,3]
    ent_table$pvalue[i] <- TE[[i]][["coef"]][1,4]
}
ent_table = as.data.frame(ent_table)
ent_table %>%
  ggplot(aes(x=names,y=ete)) +
  geom_point() +
  geom_errorbar(aes(ymin=ete-2*se,ymax=ete+2*se),width=0.3) +
  #facet_wrap(~Direction) +
  labs(y="Effective Transfer Entropy",x="Company Names",
       title = "Effective Transfer Entropy TO the Market + 95% intervals") +
  geom_hline(yintercept = 0,colour="red") +
  ylim(0,0.035) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ent_table_2 <- TE$coef
for (i in 1:length(TE)) {
    ent_table_2$names[i] <- names(TE)[i]
    ent_table_2$te[i] <- TE[[i]][["coef"]][2,1]
    ent_table_2$ete[i] <- TE[[i]][["coef"]][2,2]
    ent_table_2$se[i] <- TE[[i]][["coef"]][2,3]
    ent_table_2$pvalue[i] <- TE[[i]][["coef"]][2,4]
}
ent_table_2 = as.data.frame(ent_table_2)

ent_table_2 %>%
  ggplot(aes(x=names,y=ete)) +
  geom_point() +
  geom_errorbar(aes(ymin=ete-2*se,ymax=ete+2*se),width=0.3) +
  #facet_wrap(~Direction) +
  labs(y="Effective Transfer Entropy",x="Company Names",
       title = "Effective Transfer Entropy FROM The Market + 95% intervals") +
  geom_hline(yintercept = 0,colour="red") +
  ylim(0,0.035) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```





# K - means experiment

```{r}
fungible::monte(seed = 123,nvar = 2,nclus = 3,clus.size = c(1000,1000,1000),eta2 = c(0.70, 0.30))->dat
dat1<-as_tibble(dat$data)
kmeansObj <- kmeans(dat1, centers = 3)

factoextra::fviz_cluster(kmeansObj,data = dat1)
```

SIMULATE 3 Clusters
```{R}
set.seed(12345)
f1 <- rnorm(45, rep(1:3, each = 15), 0.2)
f2 <- rnorm(45, rep(c(1, 2, 1), each = 15), 0.2)
tibble(x=f1,y=f2,obs=1:45)->dat
dat %>% ggplot(aes(x=x,y=y)) + geom_point(colour='pink') + geom_label(aes(label=obs))

```

H CLUSTERING 

```{r}
hClustering <- dat %>% dist %>% hclust(method = 'single')

ggdendrogram(hClustering)
```

## Experimental Clusters

```{r}
f2c3<-monte(seed = 123,nvar = 2,nclus = 3,clus.size = c(1000,1000,1000),eta2 = c(0.9,0.9))[['data']] %>% as.tibble()
f2c3 %>% ggplot(aes(x=V2,y=V3)) + geom_point(colour='pink')


```
Elbow Method
```{r}
### Elbow method (look at the knee)
# Elbow method for kmeans
f3c3<-monte(seed = 123,nvar = 3,nclus = 3,clus.size = c(100,100,100),eta2 = c(0.8,0.8,0.8))[['data']] %>% as.tibble()
fviz_nbclust(f3c3, kmeans, method = "wss")
```

Gap Statistic
```{r}
fviz_nbclust(f3c3, kmeans, method = "gap_stat")
```






Silhouette 
```{r}
fviz_nbclust(f3c3, kmeans, method = "silhouette")
```


```{r,echo=TRUE, fig.show = "hold", out.width="50%"}
factors<-read.csv("dailyfactors.csv")

factors_scaled<-scale(factors[,-1]) %>% as.tibble()
factors_scaled %>%
  ggplot(aes(x=hml,y=umd)) + geom_point() + labs(title = "Cluster Plot",y='Momentum=Up minus Down', x="Value=High minus Low")

factors_scaled %>% select(hml,umd)->f2
kmeans(f2,centers=5)->kmobj
factoextra::fviz_cluster(kmobj,data = f2) + labs(title = "Cluster Plot - (KMeans Clusters)", y='Momentum=Up minus Down', x="Value=High minus Low")
```




# Algorithms and Inference

```{r}
set.seed(123)
S1=100; t=100; Price=vector(length = t)
noise=runif(t,min = -5,max = 5)
for (tt in 1:t) {
  if (tt==1) {
    Price[tt]=S1
  }
  else{
  Price[tt] = Price[tt-1] + noise[tt]
  }
}
sum(Price)/length(Price)

```


```{r}
standard_error<- function(x){
  dx=(x-mean(x))^2
  denom=length(x)*(length(x)-1)
  sumsq=sum(dx)/denom
  sqrt(sumsq)
}
standard_error(Price)

```





```{r}
factor<-Price + runif(t,1,8)^2
tibble(Price,factor)->df 
df %>% ggplot(aes(y=Price,x=factor)) + geom_point() +
  geom_smooth(method = "lm") #<< adds least squares line with standard errors

```




```{r}
df %>% ggplot(aes(y=Price,x=factor)) + geom_point() +
  geom_smooth(method = "loess") #

```







```{r}
set.seed(1235)
res<-tibble(testno=1:20, pvalue=1)
for (i in 1:20) {
  Up=rnorm(1000)
  NoUp=rnorm(1000)
  res[i,2]<-t.test(Up,NoUp)['p.value']
}

```




```{r}
library(tidyquant)
library(parameters)
uk_factors <- daily_factors
uk_factors %>%
  tq_transmute(select =hml,mutate_fun = monthlyReturn, col_rename = "hml_monthly") -> value_monthly
mean(value_monthly$hml_monthly)

standard_error(value_monthly$hml_monthly)

```

# 7. Explainable AI


```{r}
library(knitr)

popn<-as.numeric(ati::daily_factors$hml)
alpha <- 0.05
N <- 100 # sample size
m <- 1000 # number of strategies
p0<-0.99 # 99% of strategies are not profitable
m0 <- m*p0 # number of truly unprofitable strategies
m1 <- m-m0 # number of truly profitable strategies
nullHypothesis <- c( rep(TRUE,m0), rep(FALSE,m1))
delta <- 10 # average profitability of the strategies is 10%
set.seed(12)
calls <- sapply(1:m, function(i){
  control <- sample(popn,N)
  treatment <- sample(popn,N)
  if(!nullHypothesis[i]) treatment <- treatment + delta
  ifelse( t.test(treatment,control)$p.value < alpha, 
          "Called Significant",
          "Not Called Significant")
})

null_hypothesis <- factor(nullHypothesis, levels=c("TRUE","FALSE"))
table(null_hypothesis,calls) %>% kable(col.names = ) %>% kableExtra::kable_material()

```

## Substitiution Effect

```{r}
library(reticulate)
use_condaenv("r-reticulate")
```





```{python}
def getTestData(n_features=100,n_informative=25,n_redundant=25,n_samples=10000,random_state=0,sigmaStd=.0):
    from sklearn.datasets import make_classification
    np.random.seed(random_state)
    X,y=make_classification(n_samples=n_samples,
      n_features=n_features-n_redundant,
      n_informative=n_informative,n_redundant=0,shuffle=False,
      random_state=random_state)
    cols=['Informative_'+ str(i) for i in range(n_informative)]
    cols+=['Noise_'+ str(i) for i in range(n_features-n_informative-n_redundant)]
    X,y=pd.DataFrame(X,columns=cols),pd.Series(y)
    i=np.random.choice(range(n_informative),size=n_redundant)
    for k,j in enumerate(i):
        X["Redundant_"+str(k)]=X['Informative_'+str(j)]+np.random.normal(size=X.shape[0])*sigmaStd
    return X,y
    


```

```{python}
#import numpy as np
#import pandas as pd
#import statsmodels.discrete.discrete_model as sm
#X,y=getTestData(40,5,30,10000,sigmaStd=.1)
#ols=sm.Logit(y,X).fit()

#pvalues=ols.pvalues

```




```{r}

#install.packages("correlationfunnel")
library(correlationfunnel)
funnel_returns_ggplot <- dat %>%
  select(-Date,-Sector) %>%
  drop_na() %>%
    correlate(Return) %>%
    plot_correlation_funnel()

```




```{r}
#install.packages("h2o")
library(h2o)
h2o.shutdown()
h2o.no_progress()
h2o.init()
dat.h2o <- as.h2o(df_w)
set.seed(12345)
splitss <- h2o.splitFrame(dat.h2o, ratios = c(.7, .15), destination_frames = c("train","valid","test"))
names(splitss) <- c("train","valid","test")
```




```{r}
getExpectedMaxSR<-function(nTrails,meanSR,stdSR){
  # Expected Max SR controlling for SBuMT
  emc=0.577215664901532860606512090082402431042159336
  sr0=(1-emc)*qnorm(p=1-1./nTrails)+emc*qnorm(1-(nTrails*exp(1))^(-1))
  sr0=meanSR+stdSR*sr0
  return(sr0)
}
```




```{r}
getDistMaxSR<-function(nSims,nTrails,meanSR,stdSR){
  out=tibble("Max{SR}"=NA,"nTrails"=NA)
  for (nTrails_ in nTrails) {
    #1) Simulated Sharpe Ratios
    set.seed(nTrails_)
    sr<-array(rnorm(nSims*nTrails_),dim = c(nSims,nTrails_))
    sr<-apply(sr,1,scale) # demean and scale
    sr= meanSR+sr*stdSR
    #2) Store output
    out<-out %>% bind_rows(
      tibble("Max{SR}"=apply(sr,2,max),"nTrails"=nTrails_))
  }
  return(out)
}

```




```{r}
#install.packages("pracma")
library(pracma)
# Create a sequential on the log-linear scale
nTrails<-as.integer(logspace(1,4,100)) %>% unique()
plot(nTrails)
sr0=array(dim = length(nTrails))
for (i in seq_along(nTrails)) {
  sr0[i]<-getExpectedMaxSR(nTrails[i],meanSR = 0, stdSR = 1)
}
sr1=getDistMaxSR(nSims = 1000,nTrails = nTrails,meanSR = 0,stdSR = 1)

```




```{r}
library(pracma)
# Create a sequential on the log-linear scale
nTrails<-as.integer(logspace(1,4,100)) %>% unique()
plot(nTrails)
sr0=array(dim = length(nTrails))
for (i in seq_along(nTrails)) {
  sr0[i]<-getExpectedMaxSR(nTrails[i],meanSR = 0, stdSR = 1)
}
sr1=getDistMaxSR(nSims = 1000,nTrails = nTrails,meanSR = 0,stdSR = 1)

```



## Most important plot in quant finance
```{r}
sr1 %>% ggplot(aes(x=nTrails,y=`Max{SR}`)) + 
  geom_point(alpha=0.01,colour="blue") +
  geom_line(data = tibble(nTrails,`E{maxSR}`=sr0),aes(y=`E{maxSR}`, x=nTrails),colour='red') +
  labs(y="E{maxSR},MaxSR",x="Number of Trails",title="max(SR) for uninformed strategies for std(SR)=1")
```

### Inference from plot
- The experiment compares the empirical (Monte Carlo) estimate of Maximum Sharpe ratio under the null of a false strategy to that implied by the FS theorem
- The plot shows the output of the experiment for 1 to 10,000 trails.
- The code sets $V[\hat{SR_k}]=1$ and simulates the maximum Sharpe ratio 500 times, to derive a distribution of maximum Sharpe ratios for any k (number of trails).
- the y axis shows the distribution of the $max_k(\hat{SR_k})$ and the Expect  
- this results is profound, after only 100 independent backtests the expected maximum Sharpe ratio is 3.2, even when the true Sharpe ratio is zero.
- The reason is **Backtest overfitting**: when selection bias (picking the best results) takes place under multiple testing (running many alternative configurations) that backtests are likely to be false discoveries.



```{r}

```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```






























